ProbabilityBoundsAnalysis.jl:\\ an arithmetic of probabilities and sets of probabilities

### 2021-09-25 paper/paper.tex
adds more to section 2

@@

That is, in the region $[p_{i}, p_{i+1})$, $u[i]$ is a lower bound on $\overline{F}^{-1}$ and $d[i]$ is an upper bound on $\underline{F}^{-1}$. Note that the definition of the regions $[p_{i}, p_{i+1})$ or $(p_{i}, p_{i+1}]$ doesn't matter, as long as the choice is consistent. In the case that the p-box is a precise distribution and $\overline{F}^{-1}(p) = \underline{F}^{-1}(p)$ gives:  $u[i+1] = d[i]$. Figure \ref{fig:figure2} shows examples of a distribution and a p-box represented this way.

@@

\iffalse
\begin{lstlisting}[language = Julia]
  using Distributions

  function discretise(Dist, params)

    i =

  end
\end{lstlisting}
\fi

@@

\subsubsection{Operations involving intervals and distributions} %\hfill \break

\subsubsection{When dependencies between distributions are uknown} %\hfill \break

@@

\subsubsection{Inferential methods where data is limited or bad} \hfill \break

@@

An example is robust bayesian analysis [reference], where the analyst does not have enough information to reliably establish a single prior and likelihood function, but instead has a collection of priors and likelihoods, leading to a set of posteriors distributions, to which a p-box can be identified.

@@

\section{P-box arithmetic}

@@

>>>>

\usepackage[labelformat=simple]{subfig}

\renewcommand\thesubfigure{\arabic{subfigure}}
\renewcommand{\subfigurename}{Figure}

@@

That is, in the region $[p_{i}, p_{i+1})$, $u[i]$ is a lower bound on $\overline{F}^{-1}$ and $d[i]$ is an upper bound on $\underline{F}^{-1}$. Note that the definition of the regions $[p_{i}, p_{i+1})$ or $(p_{i}, p_{i+1}]$ doesn't matter, as long as the choice is consistent. In the case that the p-box is a precise distribution and $\overline{F}^{-1}(p) = \underline{F}^{-1}(p)$ gives:  $u[i+1] = d[i]$. Figure \ref{fig:figure2} shows examples of a distribution and a p-box with various outer representations.

@@

Other distributional p-boxes are constructed in a similar way. A evaluation of the cdf in terms of this representation is as follows:

\begin{lstlisting}[language = Julia]
  function cdf(F :: pbox, x :: Real)

    u = F.u; d = F.d; n = length(u);

    cdf_u = 1 - sum(x .< u)/n;
    cdf_d = 1 - sum(x .<= d)/n;

    return interval(cdf_d, cdf_u)
  end

A horizontal cut (an evaluation of the inverse cdf) and a sample of a p-box can be found as follows:

\begin{lstlisting}[language = Julia]
  function cut(F :: pbox, p :: Real)

    u = F.u; d = F.d; n = length(u);
    ps = range(0, 1, length = n+1);

    ind_u = findlast(ps[1:end-1] .<= p)
    ind_d = findfirst(ps[2:end] .>= p)

    return interval(u[ind_u], d[ind_d])
  end

  # a random sample is a random cut
  rand(F :: pbox, N :: Int) = cut.(F, rand(N))

And finally the probability measure, which we call the \textit{mass} for short:

\begin{lstlisting}[language = Julia]
  function mass(F :: pbox, x :: Interval)

    cdfHi = cdf(F, x.hi)  
    cdfLo = cdf(F, x.lo)  # returns intervals

    ub = cdfHi.hi - cdfLo.lo
    lb = max(0, cdfHi.lo - cdfLo.hi)

    return interval(lb, ub)
  end

Other available functions on p-boxes which we do not describe here are:

\begin{itemize}
  \item \textbf{env} or $\cup$: takes the envelope or hull of two p-boxes. A convex set-union of the probability sets
  \item \textbf{imp} or $\cap$: imposition or intersection of two p-boxes. Set-intersection of probability sets
  \item \textbf{subseteq} or $\subseteq$: checks if a p-box is a subset or equal to another
  \item \textbf{in} or $\in$: checks if a distribution is a member of a p-box
  \item \textbf{mixture}: a stochastic mixture of p-boxes.
\end{itemize}

@@

\subsubsection{Operations involving intervals and distributions} \hfill \break

An arithmetic operation involving an interval and a precise distribution will produce set of distributions, i.e. a unique distribution for every real number in the interval, whose p-box bounds can be found. The following function performs a binary operation between a p-box and an interval:

\begin{lstlisting}[language = Julia]
  using MomentArithmetic

  function binary(F :: pbox, X :: Interval, op)

    u_z = op.(Interval.(F.u), X)
    d_z = op.(Interval.(F.d), X)

    u_z_ = getfield.(u_z, :lo)  # get right edges
    d_z_ = getfield.(d_z, :hi)  # get left edges

    # for moments
    F_range = interval(F.u[1], F.d[end])
    F_moms = Moments(F.m, F.v, F_range);
    Z_moms = op(F_moms, X);

    return pbox(u_z_, d_z_, Z_moms.mean, 
    Z_moms.var, F.shape)
  end
\end{lstlisting}

How the moment transformations are performed are not described here, however are available in \textit{MomentArithmetic.jl} \cite{ferson2021distribution}. Figure \ref{fig:figure4} shows an example of an operation between a precise distribution and an interval.

\begin{figure}[htp]

  \centering
  \includegraphics[width=0.12\textwidth]{../examples/JuliaCon/fig4/fig4_dist.pdf}
  \raisebox{8.0mm}{\noindent\Large*}
  \includegraphics[width=0.12\textwidth]{../examples/JuliaCon/fig4/fig4_in.pdf}
  \raisebox{9.0mm}{{\Large$\rightarrow$}}
  \includegraphics[width=0.12\textwidth]{../examples/JuliaCon/fig4/fig4_pbox.pdf}

  \caption{Illustration of a p-box resulting from a binary operation between a distribution and an interval}
  \label{fig:figure4}

\end{figure}

\subsubsection{When dependencies between distributions are unknown} \hfill \break

In the case where a binary operation is performed between two distributions but whose dependence or correlation is unknown, then a single output distribution also cannot be defined. However a bounding p-box can be computed. This is the original question of Kolmogorov as described in the introduction. We call such operations between p-boxes where the marginals are fixed but no joint information is known \textit{Fréchet} operations, after the mathematician Maurice R. Fréchet, who first discovered how to bound joint distributions with fixed marginals \cite{frechet1935generalisation}. How operations between p-boxes with partial or no dependence information is known is described in section \ref{sec:pboxBinary}. Figure \ref{fig:figure5} illustrates a Fréchet operation between a uniform and a normal distribution.

\begin{figure}[htp]

  \centering
  \includegraphics[width=0.12\textwidth]{../examples/JuliaCon/fig5/fig5_dist1.pdf}
  \raisebox{8.0mm}{\noindent\Large+\small$_{\text{Fr{\'e}chet}}$}
  \includegraphics[width=0.12\textwidth]{../examples/JuliaCon/fig5/fig5_dist2.pdf}
  \raisebox{9.0mm}{{\Large$\rightarrow$}}
  \includegraphics[width=0.12\textwidth]{../examples/JuliaCon/fig5/fig5_pbox.pdf}

  \caption{Illustration of a p-box resulting from a Fréchet operation}
  \label{fig:figure5}

\end{figure}

@@

\subsubsection{Inferential methods when data is limited or bad} \hfill \break

@@

An example is robust bayesian analysis, where the analyst does not have enough information to reliably establish a single prior and likelihood function, but instead has a collection of priors and likelihoods, leading to a set of posteriors distributions, to which a p-box can be identified.

A method which is growing in popularity are inferential models \cite{martin2015inferential}. These generalise classical inference methods to account for limited statistical knowledge, and gives a statistical inference theory which is non-additive (as opposed to classical probability which is additive). These inference procedures, often based on random sets, give certain validity and performance guarantees. Inferential models can viewed from the perspective of bounding sets of distributions, and is closely linked to possibility theory \cite{liu2020inferential}. A p-box can be prescribed to such structures, leading to structures like confidence boxes or c-boxes \cite{fersoncomputing}, which follow the same arithmetic rules as p-boxes.

@@

\section{P-box arithmetic}
\label{sec:pboxarithmetic}

@@

\subsection{Binary operations}
\label{sec:pboxBinary}
